Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-20 doc/mc-paper.tex
new Lean section
Tweaked box placements
Rerrange appendix

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
\label{prop:ring:eval}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any weight function $w$:
\begin{eqnarray}
%% V1
\textstyle
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Pand_{i=1 \leq i \leq k} \phi_i,\; w \right) &=& \prod_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Por_{i=1 \leq i \leq k} \phi_i,\; w\right) &=& \sum_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:sum}
%% V2
%\rep(\boolnot \phi, w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
%\rep\left(\phi_1, w \pand \dots \pand \phi_k, w \right) &=& \prod_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\phi_1, w \por \dots \por \phi_k, w \right) &=& \sum_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:sum}
%\rep\left(\Pand_{i=1 \leq i \leq k} \phi_i, w \right) &=& \prod_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\Por_{i=1 \leq i \leq k} \phi_i, w\right) &=& \sum_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}
As is described in Section~\ref{subsection:counting}, we have proved these three equations using \lean{}.

@@

\begin{minipage}{0.62\textwidth}

@@

\begin{minipage}{0.45\textwidth}

@@

Our experiments are designed to evaluate the following:
\begin{itemize}
\item
  whether or toolchain can handle challenging benchmark problems,
\item  
the effectiveness of some design choices and optimizations,
\item
how the performance of our toolchain compares to that of the MICE model counter verifier,
\item
the tradeoffs between the prototype and verified checker.  These experiments are described in Section~\ref{app:lean:evaluation}.
\end{itemize}

@@

>>>>

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
\label{prop:ring:eval}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any weight function $w$:
\begin{eqnarray}
%% V1
\textstyle
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Pand_{i=1 \leq i \leq k} \phi_i,\; w \right) &=& \prod_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Por_{i=1 \leq i \leq k} \phi_i,\; w\right) &=& \sum_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:sum}
%% V2
%\rep(\boolnot \phi, w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
%\rep\left(\phi_1, w \pand \dots \pand \phi_k, w \right) &=& \prod_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\phi_1, w \por \dots \por \phi_k, w \right) &=& \sum_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:sum}
%\rep\left(\Pand_{i=1 \leq i \leq k} \phi_i, w \right) &=& \prod_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\Por_{i=1 \leq i \leq k} \phi_i, w\right) &=& \sum_{i=1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}
As is described in \cref{section:formally-verified-toolchain}, we have proved these three equations using \lean{}.

@@

\begin{minipage}{0.58\textwidth}

@@

\begin{minipage}{0.42\textwidth}

@@

Our experiments are designed to evaluate the following:
\begin{itemize}
\item
  whether or toolchain can handle challenging benchmark problems,
\item  
the effectiveness of some design choices and optimizations,
\item
the comparative performance of the prototype and verified checkers, and
\item
how the performance of our toolchain compares to that of the MICE model counter verifier.
\end{itemize}

@@

\subsection{Evaluation of the Verified Checker}
\label{app:lean:evaluation}

\input{figure-check}

We ran the \lean{} checker on the 80 benchmark problems for which the
generated CRAT file had at most 10~million total clauses.  These are the
same benchmarks used in evaluating the optimizations (Section
\ref{app:experiment:optimize}).
All of the checks completed and confirmed the outcome of the prototype checker.

Figure~\ref{fig:lean} shows the
performance of both the \lean{} checker and the prototype
checker on the Y axis, as functions of the total number of clauses on the X axis.
As would be expected, the \lean{} checker consistently runs slower
than the prototype checker.  
The ratio of the runtime for \lean{} versus the runtime for
the prototype checker had a harmonic mean of $5.94$.

Importantly, however, it can be seen that both checkers show the same
overall performance trends.  The ratio of the two runtimes was less
than $8.0$ for all but three small benchmarks.  This seems like a
reasonable price to pay for a rigorous guarantee of correctness, and
we are confident that we can reduce this gap with more effort in
optimizing the verified checker.
